---
title: "MadStork - Portfolio Optimization Template"
author: "Haarstick"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  html_document:
    toc: true
    theme: united
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, error = TRUE, eval=FALSE,
                      warning = FALSE, message = FALSE,
                      fig.align = "center", fig.width = 10, fig.height = 5)
t1 <- Sys.time()
```


```{r r-config}

# Library Path
lib_path <- .libPaths()[1]

# R packages
packages <- c("a2munge", "a2modeler", "a2charter", "dplyr", "tidyr", "readxl", "tibble", "purrr", "ggplot2", "forcats", "stringr", "lubridate", "rlang", "madstork", "madstork.opt", "here", "knitr", "DT", "ggrepel")


```


```{r r-setup}

# Load Libraries
for(pck in packages) library(pck, lib.loc = lib_path, character.only = TRUE)

```



```{r portfolio-config}

# Path to MadStork Portfolio Object
portfolio_path <- ""
```



```{r estimates-config}

# Path to Estimates Object
estimates_path <- ""
```


```{r load-objects}

# Load portfolio object
port <- load_portfolio(portfolio_path)

# load market predictions
est <- readRDS(estimates_path)

```


```{r port-stats}

# Calculate Portfolio Stats
port_stats <- get_estimated_port_stats(pobj = port, eobj = est, port_only = FALSE)
kable(port_stats,
      digits = 3,
      caption = "Current Portfolio Statistics")
```


```{r constraints}

# Define Constraints
const <- constraints(symbols = est$symbols) %>%
  add_symbol_constraint(min = 0.0, max = 0.20) %>% 
  add_cash_constraint(min = 0.02, max = 0.05)

```


```{r optimization-config}

# Set Optimization Parameters
.target <- 'sharpe'   
.amount <- 5000
.lot_size <- 1
.max_runtime <- 300
.npairs <- 20
.max_iters <- 20
.improve_lag <- 5
.min_improve <- .001

```


```{r portfolio-optimization}

# Prices
prices <- get_current_prices(est$symbols, dividends = TRUE)

# Create Optimization
po <- portfolio_optimization(port, est, const, prices, target = .target)

# Optimize
po_opt <- optimize(po,
                   n_pairs = .npairs,
                   amount        = .amount,
                   lot_size      = .lot_size,
                   max_iter      = .max_iters,
                   max_runtime   = .max_runtime,
                   improve_lag   = .improve_lag,
                   min_improve   = .min_improve,
                   plot_iter     = FALSE)
```


```{r execution, eval=TRUE, ref.label=c('r-config', 'r-setup', 'portfolio-config', 'estimates-config', 'load-objects', 'constraints', 'optimization-config', 'portfolio-optimization')}

```


```{r port-print, eval=TRUE, results='asis'}
cat("##", "Portfolio:", port$name, "\n")

```



```{r port-values, eval=TRUE}
port$market_value %>%
  filter(last_updated == max(last_updated)) %>% 
  kable(.,
        digits = 2,
        format.args = list(big.mark = ","),
        caption = "Current Portfolio Market Value")
```


```{r port-stats-print, eval=TRUE, ref.label=c('port-stats')}

```


```{r port-constraints-print, eval=TRUE}

# Check Constraints 
check_constraints(const, port, est) %>% 
  datatable(.,
            caption = "Portfolio Constaints Check") %>% 
  formatRound(6, digits = 4)
```



## Portfolio Optimization

Optimization goal is to `r po_opt$criteria` portfolio __`r  po_opt$target`__

* Version: __`r po_opt$version`__  
* Created on: __`r as.character(po_opt$created_on)`__ 
* Created by: __`r po_opt$user`__
* Description: _`r po_opt$desc`_  


### Estimates

```{r esimates-chart, eval = TRUE}

get_estimates_stats(est) %>% 
  ggplot(., aes(y=mu, x=sd)) + 
  geom_point(color = "red", size = 2) +
  geom_text_repel() + 
  geom_abline(linetype = 5, color = "grey75") +
  theme_sncr() +
  guides(color = FALSE, label = FALSE) +
  scale_color_madstork() +
  scale_x_continuous(labels = scales::percent) +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Estimates Frontier Plot",
       subtitle = "",
       x = "Std Dev",
       y = "Return")


get_estimates_stats(est) %>% 
  mutate(yield = yield / 100) %>% 
  datatable(.,
            filter = "top",
            caption = "Portfolio Optimization Estimates") %>% 
  formatPercentage(c(2, 3, 5), digits = 1) %>% 
  formatRound(4, digits = 3)
```


### Constraints

The following constraints were placed on the optimization

```{r constraint-print, eval=TRUE}
print(const)
```


### Parameters

The optimization parameters were set as follows 

```{r optimization-config-print, eval=FALSE, echo=TRUE, ref.label='optimization-config'}

```


### Results


```{r po-stats, eval=TRUE}

# Estimated Optimal Port Stats
get_estimated_port_stats(po_opt$optimal_portfolio, po_opt$estimates) %>% 
  kable(.,
        digits = 3,
        caption = "Optimal Portfolio Statistics")
```


```{r po-values, eval=TRUE}

# Optimal Port Values and Market Values
po_opt$portfolio_values %>% 
  filter(iter == max(iter)) %>% 
  select(-iter) %>% 
  kable(.,
        digits = 2,
        format.args = list(big.mark = ","),
        caption = "Optimal Portfolio Values")

get_estimated_port_market_value(po_opt$optimal_portfolio, po_opt$estimates) %>% 
  kable(.,
        digits = 2,
        format.args = list(big.mark = ","),
        caption = "Estimated Optimal Portfolio Market Values")

```


```{r po-target-chart, eval=TRUE}

# Portfolio Target Path Chart
po_target_chart(po_opt)
```


#### Symbol Shares

```{r po-share-chart, eval=TRUE, fig.height=6}

# Current vs Optimal Symbols Share
map_df(list(port, po_opt$optimal_portfolio),
       ~get_symbol_estimates_share(pobj = ., eobj = po_opt$estimates), .id = "port") %>%
  spread(key = "port", value = "portfolio_share") %>% 
  select(symbol, Current = `1`, Optimal = `2`) %>% 
  datatable(.,
            caption = "Current vs Optimal Symbols Share") %>% 
  formatPercentage(-1, digits=1)

# Portfolio Optimization Share Chart
po_symbol_share_chart(po_opt)

# Share Data
map_df(po_opt$portfolios,
       ~get_symbol_estimates_share(pobj = ., eobj = po_opt$estimates), .id = "iter") %>%
  spread(key="symbol", value = "portfolio_share") %>% 
  arrange(iter) %>% 
  datatable(.,
            options = list(scrollX = TRUE),
            caption = "Symbol Share Path") %>% 
  formatPercentage(-1, digits=1)
```


#### Constraints

```{r po-constraint-chart, eval=TRUE, fig.height=10}

# Portfolio Constraints Chart
po_constraints_charts(po_opt)

# Check Constraints 
check_constraints(po_opt$constraints,
                  po_opt$optimal_portfolio,
                  po_opt$estimates) %>% 
  datatable(.,
            filter = "top",
            caption = "Optimal Portfolio Constraints Check") %>% 
  formatRound(6, digits = 4)
```


#### Trade Pairs

```{r po-trade-pairs, eval=TRUE}

# Optimization Trade Pairs
po_opt$trade_pairs %>% 
  arrange(-trades) %>% 
  datatable(.,
             filter = "top",
            caption = "Optimization Trade Pairs") %>% 
  formatRound(4:6, digits = 3)
```


#### Trades


```{r po-trades, eval=TRUE}

# Optimal Trades
po_opt$trades %>% 
  datatable(.,
            filter = "top",
            extensions = 'Buttons', 
            options = list(
              dom = 'Bfrtip',
              buttons = c('copy', 'csv', 'print')),
            caption = "Potfolio Optimization Trades") %>% 
  formatCurrency("amount") %>% 
  formatRound("price", digits = 2)

```


## Appendix

```{r config, echo=TRUE, eval=FALSE, ref.label=c('r-config')}

```

* runtime: `r round(difftime(Sys.time(), t1, units = "mins"), 2)` minutes