---
title: "Market Index Forecast"
author: "Haarstick"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  html_document:
    toc: true
    theme: united
params:
  raw_data_dir: ./data/raw
  predictions_dir: ./data/predictions
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, error = TRUE, eval=FALSE,
                      warning = FALSE, message = FALSE,
                      fig.align = "center", fig.width = 10, fig.height = 5)
t1 <- Sys.time()
```


```{r r-config}

# Library Path
lib_path <- .libPaths()[1]

# R packages
packages <- c("mad4sight", "madutils", "madstork", "dplyr", "tidyr", "readxl", "tibble", "purrr", "ggplot2", "gridExtra", "stringr", "lubridate", "DT", "rlang", "here")


```


```{r r-setup}

# Load Libraries
for(pck in packages) library(pck, lib.loc = lib_path, character.only = TRUE)

```



```{r data-import}

# Get Shiller Market Data
dataset_name <- "Shiller CAPE"
url <- "http://www.econ.yale.edu/~shiller/data/ie_data.xls"
dataset_name <- paste0("Shiller-", as.character(Sys.Date()), ".xls")
destfile <- paste(params$raw_data_dir, dataset_name, sep="/")
download.file(url, destfile, mode = "wb")
raw <- readxl::read_excel(path = destfile, 
                          sheet = "Data",
                          skip = 7,
                          col_names = TRUE,
                          col_types = c("text", rep("numeric", 10)))
#file.remove(destfile)
```


```{r data-prep}

# Prep data for analysis
clean <- raw %>% 
  dplyr::rename_all(tolower) %>% 
  dplyr::rename(gs10 = `rate gs10`) %>% 
  tidyr::separate(., "date", into = c("year", "month"), sep = "\\.") %>%
  dplyr::mutate(month = str_sub(month, 1, 2)) %>% 
  dplyr::mutate_at(c('month', 'year'), as.numeric) %>% 
  dplyr::mutate(day = as.numeric(days_in_month(month)),
                date = paste(paste(year, month, sep = "-"), day, sep = "-"),
                date = as.Date(date)) %>% 
  dplyr::mutate(index = row_number()) %>% 
  dplyr::mutate(gs10 = gs10/100) %>% 
  dplyr::filter(! is.na(date)) %>% 
  dplyr::slice(121:n()) %>% 
  dplyr::select(index, date, price, dividend, gs10, cpi, cape)
```


```{r data-impute}

# Impute any missing values
for(var in c("dividend", "gs10", "cpi", "cape")) {
  
  # get missing values and set horizon
  horizon <- clean %>% 
    pull(!!sym(var)) %>% 
    tail(120) %>% 
    is.na() %>% 
    sum()
  
  if(horizon > 0) {
    
    # create univariate seer and replace NAs with forecasts
    var_seer <- clean %>% 
      select_at(var) %>%
      na.omit() %>% 
      seer(.,
           var,
           sampling = samples(method = "split",
                              args = list(ratio = .9)),
           models = list(model(algo = "auto.arima"),
                         model(algo = "ets")),
           selection = model_selection(),
           confidence_levels = c(80, 95),
           horizon = horizon,
           forecast_xreg = NULL,
           backend = "sequential")
    
    clean[is.na(clean[[var]]), var] <- var_seer$forecast$predicted
  }
}

```



```{r transform-pipes}

# Prep pipeline for analysis
prep_pipe <- pipeline(
  expr = function(df) {
    df %>% 
      dplyr::mutate(inf_1yr         = cpi / lag(cpi, 12) - 1,
                    gs10_real       = gs10 - inf_1yr,
                    gs10_real_d_1yr = gs10_real - lag(gs10_real, 12)) %>% 
      dplyr::select(price, cape, gs10_real_d_1yr) %>% 
      na.omit() %>% 
      dplyr::mutate_at(setdiff(colnames(.), 'price'), funs((. - mean(.)) / sd(.))) %>% 
      na.omit() 
  },
  desc = "Short Term Model data preperation pipeline. Adds features & standardizes",
  uid  = "prep-pipe"
)


# Lag pipe for created lagged predictors
lag_pipe <- pipeline(
  expr = function(df, y_var, n) {
    x_vars <- setdiff(colnames(df), y_var)
    df %>% 
      dplyr::mutate_at(x_vars, funs(lag(., n))) %>% 
      na.omit() 
  },
  desc = "Creates lagged values for predictors",
  uid  = "lag-pipe"
)
```


```{r transform-data}

# Transform Data
forecast_horizon <- 12
y_var <- "price"

df <- flow(clean, prep_pipe) %>% 
  flow(., lag_pipe, y_var = y_var, n = forecast_horizon)

```



```{r feature-pipelines}

# Cape Pipeline
cape_pipe <- pipeline(expr = function(df) {
  dplyr::select(df, price, cape)  
},
uid  = "cape-only-pipe",
desc = "Pipe with only CAPE predictor")


# base pipeline
base_pipe <- pipeline(expr = function(df) {
 dplyr::select(df, price)  
},
uid  = "base-pipe",
desc = "Pipe with no predictors")

```


```{r sampling}

# Set Sampling
sampling <- samples(method = "slices", args = list(width = 240, horizon = 12, skip = 11))

```


```{r models}

# Set models
m1 <- model(algo = "auto.arima", pipe = base_pipe, uid = "aa-base")
m2 <- model(algo = "auto.arima", pipe = cape_pipe, uid = "aa-cape")
m3 <- model(algo = "auto.arima", pipe = pipeline(expr = identity), uid = "aa-full")
m4 <- model(algo = "ets", pipe = base_pipe, uid = "ets-base")

```



```{r selection}

# set model selection
selection <- model_selection(measure = "RMSE", n = 2, weights = "weighted")
```



```{r forecast-xreg}

# Get forecast xreg
forecast_xreg <- flow(clean, prep_pipe) %>% 
  tail(forecast_horizon) %>% 
  select(-!!sym(y_var))

```



```{r seer}

# Create Seer
foreseer <- seer(df,
                 y_var = y_var,
                 sampling = sampling,
                 models = list(m1, m2, m3, m4),
                 selection = selection,
                 confidence_levels = c(80, 95),
                 horizon = forecast_horizon,
                 forecast_xreg = forecast_xreg,
                 backend = "sequential",
                 user = NULL,
                 uid = "mkt-shiller-forecast",
                 desc = "seer to forecast 1 year market returns")

```




```{r execution, eval=TRUE, ref.label=c('r-config', 'r-setup',  'data-import', 'data-prep', 'data-impute', 'transform-pipes','data-transform', 'feature-pipelines', 'sampling', 'models', 'selection', 'forecast-xreg', 'seer')}

```


```{r seer-print, eval=TRUE, results='asis'}
cat("##", "Seer Details", "\n")

```

* uid: __`r foreseer$uid` __ 
* Created on: __`r as.character(foreseer$created_at)`__ 
* Created by: __`r foreseer$user`__
* Description: _`r foreseer$desc`_  


```{r data-print, eval=TRUE, results='asis'}
cat("###", "Input Data:", dataset_name, "\n")

```

* Location: __`r destfile`__
* Type: __xls__

<br>

#### Sample Records


```{r sample-records, eval=TRUE}

# Datatable with sample records
clean %>%
  DT::datatable(.,
                options = list(scrollX = TRUE),
                caption = paste(dataset_name, "Sample Records")) %>% 
  DT::formatRound(colnames(clean), digits=3)
```


```{r target-print, eval=TRUE, results='asis'}
cat("###", "Target:", foreseer$y_var, "\n")

```


```{r target-summary, eval=TRUE}

clean %>% 
  mutate(ret = price/lag(price, 12) - 1) %>% 
  ggplot(., aes(x=ret)) +
  geom_histogram(fill = madstork_cols()[3], alpha=.6, bins=50, color = "grey25") +
  scale_x_continuous(labels = scales::percent) +
  theme_minimal()+
  labs(title = "US Stock Market Yearly Return Distribution",
       x = "Yearly Return")

  
clean %>% 
  mutate(ret = price/lag(price, 12) - 1) %>% 
  dplyr::summarise_at("ret", funs(min, mean, sd, max), na.rm=TRUE) %>% 
  knitr::kable(., 
               digits = 3,
               caption = "US Stock Market Monthly Return Summary")
```

<br>


### Features

```{r feature-plots, eval=TRUE, fig.height=8}

clean %>% 
  filter(date >= max(date) - years(10)) %>% 
  gather(key = "variable", value = "value", -index, -date, -!!sym(foreseer$y_var)) %>%
  ggplot(., aes(x=date, y=value)) +
  geom_line(color = madstork_cols()[3])+
  facet_wrap(~variable, scales = "free", ncol=2)+
  theme_minimal()+
  labs("Shiller Market Covariates")
```


<br>

### Pipelines


```{r pipeline-print, eval=TRUE}

# base pipeline
base_pipe

# cape only
cape_pipe

```

<br>

### Models


```{r models-print, eval=TRUE, results='asis'}

for(i in seq_along(foreseer$models)) {
  cat("####", foreseer$models[[i]]$uid)
  cat("\n")
  cat("* Algorithm:", paste0("__", foreseer$models[[i]]$algo, "__"), "\n")
  cat("* Pipeline:", paste0("__", foreseer$models[[i]]$pipe$uid, "__"), "\n")
  cat("\n")
  
  cat("##### Parameter Tuning:\n")
  for(p in seq_along(foreseer$models[[i]]$args)) {
    cat("*",
        paste0(names(foreseer$models[[i]]$args)[p], ":"),
        paste(foreseer$models[[i]]$args[[p]], collapse = ", "),
        "\n")
  }
  cat("\n")
}

```

<br>

### Leaderboard


```{r leaderboard, eval=TRUE}
foreseer$performance %>% 
 filter(sample == "validation") %>% 
  dplyr::arrange_at(foreseer$selection$measure) %>% 
  knitr::kable(.,
               digits = 3,
               caption = "Seer Model Leaderboard")
```

<br>



### Model Validation Predictions



```{r val-preds, eval=TRUE}


val_preds <- foreseer$predictions %>% 
  filter(sample == "validation") %>% 
  unnest() %>% 
  inner_join(
    df %>% 
      mutate(date = max(clean$date) - months((n()-1):0), 
             rn = row_number()),
    by = "rn"
  ) %>% 
  mutate(resid = !!sym(y_var) - predicted) %>% 
  select(rn, index, uid, date,!!sym(y_var), predicted, resid) %>%
  arrange(uid, rn)
```


```{r val-preds1}

plt1 <- ggplot(val_preds, aes(x=predicted, y=!!sym(y_var))) +
  geom_point(alpha=.4)+
  geom_smooth(aes(color = uid), alpha=.2, se=FALSE, method="lm")+
  guides(color = FALSE) +
  theme_minimal()+
  labs(title = "Predicted vs Actual",
       y = "1yr Market Return",
       x = "Predicted")

plt2 <- ggplot(val_preds, aes(x=predicted, y=resid)) +
  geom_point(alpha=.4)+
  geom_smooth(aes(color = uid), alpha=.2, se=FALSE, method="lm")+
  theme_minimal()+
  labs(title = "Predicted vs Residual",
       y = "Residual",
       x = "Predicted")
 
 gridExtra::grid.arrange(plt1, plt2, ncol=2)

```




#### Trailing 5 years


```{r, val-preds-5yr}

val_preds_5yr <- filter(val_preds, date > max(date, na.rm = TRUE) - years(5))

ggplot(val_preds_5yr, aes(y=predicted, x=date, color = uid)) +
  geom_line(aes(y=predicted)) +
  geom_point(aes(y=predicted)) +
  geom_line(data = filter(val_preds_5yr, uid == unique(uid)[1]),
            aes(y=!!sym(y_var)), color = "black", linetype=5, size=1.1) +
  facet_wrap(~index, ncol = 5, scales = "free_x") +
  scale_color_madstork() +
  theme_minimal() +
  theme(legend.position = "bottom", legend.direction = "horizontal") +
  labs(title = "Predicted vs Actual",
       subtitle = "Trailing 5 years")

```


#### Financial Downturn


```{r, val-preds-crash}

val_preds_crash <- filter(val_preds, year(date) %in% c(2007, 2008))

ggplot(val_preds_crash, aes(y=predicted, x=date, color = uid)) +
  geom_line(aes(y=predicted)) +
  geom_line(data = filter(val_preds_crash, uid == unique(uid)[1]),
            aes(y=!!sym(y_var)), color = "black", linetype=5, size=1.1) +
  facet_wrap(~index, ncol = 5, scales = "free_x") +
  scale_color_madstork() +
  theme_minimal() +
  labs(title = "Predicted vs Actual",
       subtitle = "Financial Downtown of 2008")

```

### Final Model


```{r final-model, eval=TRUE, results='asis', fig.align='center'}

foreseer$final_model %>% 
  select(uid, forecast_wt) %>% 
  knitr::kable(.,
               table.attr = "style='width:30%;'",
               caption = "Final Model Parameters")

```


```{r final-model-coef, eval=TRUE}

# Final Model Coefs
foreseer$final_model$fit %>%
  map(coef)
```


<br>

### Forecasts

```{r forecast-xreg-print, eval=TRUE}

knitr::kable(forecast_xreg,
               digits = 3,
               caption = "Forecast Covariates")
```

<br>

```{r predictions-plot, eval=TRUE}

clean %>% 
  select(date, price) %>% 
  tail(60) %>% 
  bind_rows(foreseer$forecast %>% 
              select(-index) %>% 
              mutate(date = max(clean$date) + months(1:forecast_horizon))) %>% 
ggplot(., aes(x=date))+
  geom_line(aes(y=!!sym(y_var)), color = "black", size=1.05)+
  geom_ribbon(aes(ymin=low95, ymax=high95), fill = "lightblue", alpha=.4) +
  geom_ribbon(aes(ymin=low80, ymax=high80), fill = "lightblue", alpha=.6) + 
  geom_line(aes(y=predicted), color = "dodgerblue", size=1.05)+
  theme_minimal() +
  labs(title = "Market 12 Month Forecast",
       y = "SP500 Price")


forecasts <- foreseer$forecast %>% 
  mutate(date = max(clean$date) + months(1:forecast_horizon),
         return = predicted / na.omit(c(tail(clean$price, 1), lag(predicted, 1))) - 1,
         cum_return =  predicted / tail(clean$price, 1) - 1) %>% 
  select(date, return, cum_return, predicted, low80, high80, low95, high95) 

knitr::kable(forecasts,
             digits = 3,
             caption = "Market 12 Month Forecast")
```


```{r, long-term-forecasts, eval=TRUE}

# Long term forecasts
n <- 120
train_df <- clean %>% 
  mutate(ret = lead(price, n)/price - 1,
         div_prem = (dividend/price) - gs10) %>% 
  select(ret, cape, div_prem) %>% 
  mutate_at(setdiff(colnames(.), "ret"), funs((. - mean(., na.rm = TRUE))/sd(., na.rm = TRUE))) %>% 
  mutate(sample = row_number() %% n + 1) 

train_df %>%
  base::split(.$sample) %>% 
  map(~lm(ret~., data = .)) %>% 
  map_df(~as.numeric(predict(., newdata = tail(train_df, 1)))) %>% 
  gather(key = "sample", value = "prediction") %>% 
  summarise_at(c("prediction"), funs(min, mean, sd, max)) %>% 
  knitr::kable(.,
             digits = 3,
             caption = "Market Long Term 10 Year Forecast")

```




```{r predictions-write, eval=TRUE, echo=FALSE}

file_name <- paste(c("market-forecast-", as.character(Sys.Date()), ".csv"), collapse = "")
current_pred_path <- paste(params$predictions_dir, file_name, sep ="/")
write.csv(forecasts, current_pred_path, row.names = FALSE)
```

* Current predictions written to file here: _`r current_pred_path`_

<br>

## Configuration

```{r regressor-config, echo=TRUE, eval=FALSE, ref.label=c('feature-pipelines', 'sampling', 'models', 'selection', 'seer')}

```


## Appendix

```{r config, echo=TRUE, eval=FALSE, ref.label=c('r-config')}

```


* Created at: `r Sys.time()`
* runtime: `r paste(round(difftime(Sys.time(), t1), 2), "minutes")`